---
title: "GLMs_Practical_3"
author: "Zane Hassoun"
date: "11/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# EIA <- read.csv('Documents/EIA.csv')
EIA$impact <- as.factor(EIA$impact)
EIA$MonthOfYear <- as.factor(EIA$MonthOfYear)
EIA$Year <- as.factor(EIA$Year)
attach(EIA)
library(car)
```
## Question 1: 
If the linear predictor for the fit.poisSqrt model is:
ηit = β0 + β1x1it + ... + β26x26it
and the coefficients are listed in the order of the output produced using the code above, which of the
following describes the fit.poisSqrt model when the tide is ’Slack’, in Month 10 and Year 10?

Answer: A) skips from 4 to 13 to 16

## Question 2: Which of the following BEST fits the model 

Answer: A) its a sqrt link not log link or log sqrt or exp
```{r}
#Fitting the model
fit.poisSqrt <- glm(count ~ tidestate + observationhour + DayOfMonth + MonthOfYear +
Year + x.pos + y.pos + Year:x.pos + Year:y.pos, data = EIA,
family = poisson(link = 'sqrt'))
#Getting a summary of the dataset to evaluate
summary(fit.poisSqrt)
```

## Question 3 : If the linear predictor for the fit.pois model is:
ηit = β0 + β1x1it + ... + β26x26it
and the coefficients are listed in the order of the output produced using the code above, which of the
following describes the fit.pois model when the tide is ’Flood’, in Month 1 and Year 12?

Answer: A) Because it fits the numbers right

Question 4: Which of the following BEST describes the fit.pois model?

Answer: e) its a log model so to get it back exp of the log is your lambda. 

```{r}
#Fitting a new model link now is log link

fit.pois <- glm(count ~ tidestate + observationhour + DayOfMonth + MonthOfYear + Year +
x.pos + y.pos + Year:x.pos + Year:y.pos, data = EIA, family = poisson)
#summarising the data

x = 0:(length(fit.pois$coefficients)-1)
View(print(data.frame(x,fit.pois$coefficients)))
```

Instruction: 
Refit the Poisson model with a log link and an offset term (using area as an effort term) and check for
collinearity using the VIF function

Question 5 5. Which of the following about collinearity is FALSE? [1]
Answer: A) I think but not sure could be  e) ask erik

```{r}
fit.pois_off <- glm(count ~ tidestate + observationhour + DayOfMonth + MonthOfYear + Year +
x.pos + y.pos + Year:x.pos + Year:y.pos, data = EIA, offset=log(area), family = poisson)

summary(fit.pois_off)
vif(fit.pois_off)
```
Instructions: 
Compare the AIC scores for the fit.pois, fit.poisSqrt, and fit.poisOff models.

Question 6 Which of the following about these AIC results is FALSE? [1]

Answer: e) I think but not sure
```{r}
AIC(fit.pois, fit.poisSqrt, fit.pois_off)
```


Instructions: 
Carry out automated stepwise selection on the fit.poisOff governed by the AIC and BIC criteria:
call these new models step.poisOff and step.poisOff_BIC respectively (as before set direction =
"both").
Question 7: Based on these stepwise-selection results, which of the following is FALSE?

Answer: d) only two categories for tidestate
```{r}
#Step function for AIC
small_step_AIC = step(fit.pois_off, direction = "both")
step.poisOFF = glm(count ~ tidestate + observationhour + MonthOfYear + Year + x.pos + 
    y.pos + Year:x.pos + Year:y.pos, data = EIA, offset=log(area), family = poisson)
```

```{r}
#Step Function for BIC
small_step_BIC = step(fit.pois_off, direction = "both", k = log(nrow(EIA)))
step.poisOff_BIC = glm(count ~ tidestate + observationhour + MonthOfYear + Year + x.pos + 
    y.pos + Year:x.pos + Year:y.pos, offset=log(area), family = poisson)
```

Instructions: Perform likelihood ratio test results using the Anova function for the step.poisOff_BIC model. 

Queston 8: Based on these results, which of the following is FALSE?

Answer: c) C is wrong anova gives us the type two tests its quite different. 

```{r}
Anova(step.poisOff_BIC)
```
Instructions:Carry out all-possible-subsets selection on the step.poisOff_BIC model using the dredge function and
the default ranking criteria. 

Question 9: Based on these results, which of the following is FALSE?

Answer: a)

```{r}
library(MuMIn)
options(na.action='na.fail')
dredged_model = dredge(step.poisOff_BIC)
dredged_model
```

Question 10: Based on the step.poisOff_BIC model results, which of the following is FALSE? 

Answer: D)
```{r}
summary(step.poisOff_BIC)
```

Question 11: Based on the step.poisOff_BIC model, what is the predicted value on the scale of the response, when
tidestate=EBB, observationhour=10, month=1, year=11, x.pos=-2061, y.pos=-1158 and the area of the
cell is the mean of the area of all cells in the EIA data set? Report your answer to 3 decimal places.

```{r}
prediction_df = data.frame(tidestate=as.factor("EBB"),
                           observationhour=10,
                           MonthOfYear = c(as.factor("1"), as.factor("5")),
                           Year = as.factor("11"),
                           x.pos=-2061, 
                           y.pos=-1158, 
                           area = mean(EIA$area))


predictions = predict.glm(step.poisOff_BIC,prediction_df)
exp(predictions)
```
Question 12: Based on the step.poisOff_BIC model, what is the ratio of the predicted numbers in month 5 compared
to predicted numbers in month 1? Report your answer to 3 decimal places.

0.81

# Section 2 of the practical Modelling Overdispersed Models

Instructions Check for overdispersion in the step.poisOff_BIC model using family=quasipoisson. Call this new model step.poisOffOD. 
Question 13: Based on these results, what is the estimated dispersion parameter? Report your
result to one decimal place

15.1102

```{r}
step.poisOffOD = glm(count ~ tidestate + observationhour + MonthOfYear + Year + x.pos + 
    y.pos + Year:x.pos + Year:y.pos, offset=log(area), family = quasipoisson())

summary(step.poisOffOD)
```


Instructions: Compare the p-values with the overdispersed model with the p-values obtained under a strictly Poisson model (when the dispersion parameter=1).

Question 14: Which of the following is FALSE?

Answer: C) 

```{r}
summary(step.poisOff_BIC)
```
```{r}
summary(step.poisOffOD)
```
15. Based on the overdispersed results, is the following statement TRUE or FALSE? 

Answer: TRUE

# Section 2.3 Model Diagnostics

Instructions: In this section we are going to:
• Assess any residual patterns using the residualPlots function in the car library
• Check for any correlation present in the residuals.
• Assess what we have learned from this modelling process with reference to our research questions.
Based on the analysis from the previous section update your model dropping insignificant covariates and/or
interaction terms.


Question 16: Using the residualPlots function in the car library in R, which of the following statements about linearity
is TRUE?

Answer: c) i think
```{r}
residualPlots(step.poisOffOD)
```

Instructions: Make a plot of observed vs fitted. 

Question 17: Which of the following statements is FALSE?

Answer: c) is false cuz we did log transform so there cant be
```{r}
library(ggplot2)
plot_dta_17 = data.frame("Observations" = EIA$count, "Fitted_Values" = step.poisOffOD$fitted.values)

ggplot(data = plot_dta_17, aes(x = Observations, y = Fitted_Values)) + 
  geom_point()+geom_abline(intercept=0,slope=1,color="red")
```


```{r}
phi_hat = summary(step.poisOffOD)$dispersion
scaled_resid =  (EIA$count-plot_dta_17$Fitted_Values)/sqrt(phi_hat*plot_dta_17$Fitted_Values)

ggplot(data = plot_dta_17, aes(x = Fitted_Values, y = scaled_resid))+
         geom_point()

```
Instructions: Use acf plots to determine the nature of any scaled residual correlation present. 

Question 19: Which of the following statements is FALSE?

Answer: C) 
```{r}
par(mfrow =c(1,2))
set.seed(5)
acf(rnorm(length(scaled_resid)), lag.max = 40, main = "Ideal correlation")
acf(scaled_resid, main = "Actual correlation")
```
Question 20: Which of the following about summarising our model is FALSE?

Answer: E)
```{r}
summary(step.poisOffOD)
```














